Consumption and resistance across states

### 2017-02-27 inequality/consumption-inequality.Rmd
updates ineq, states

normalize = function(x) x / max(x)
juran = function(x) {
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

@@

@@

@@

>>>>

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

@@

```{r inequality_hist, cache=TRUE}
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()
```

@@

## Characteristics by consumption

I broke down the beneficiaries based on their consumption: those who had no antibiotic claims were in one "tile", and I separated the rest into $n$-tiles.

```{r chars_by_consumption, cache=TRUE}
ntile_zero = function(x, n) {
  out = x
  out[x > 0] = ntile(out[x > 0], n)
  out
}

bounds_group = function(x, upper_limits) {
  upper_limits = as.integer(upper_limits)
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('<= %i', upper_limits[i])
    } else {
      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
    }
  })
  top_level = sprintf('> %i', tail(upper_limits, 1))
  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
                        levels, upper_limits) %>%
    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))

  levels = c(levels, top_level)

  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

usage %>%
  group_by(age, sex, race, chronic, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  ungroup() %>% 
  mutate(claims_group=bounds_group(n_claims, c(0, 1, 2, 5, 10, 20, 50, 100)),
         is_female=if_else(sex=='female', 1, 0),
         is_white=if_else(race=='white', 1, 0)) %>%
  group_by(claims_group) %>%
  summarize(n_beneficiaries=n(),
            mean_age=mean(age),
            percent_female=mean(is_female)*100,
            percent_white=mean(is_white)*100,
            mean_n_chronic_cond=mean(chronic),
            mean_n_claims=mean(n_claims)) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>% 
  kable
```

@@

Surprisingly, the characteristics of the beneficiaries are fairly constant by consumption $n$-tile among those who took antibiotics. Those who took no antibiotics where more male and less white.

### 2017-02-27 states/resistance/state-report.Rmd
updates ineq, states

```{r run_models}
safe_spearman = function(df, expl) {
  if (dim(df)[1] < 3L) {
    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
  } else {
    cor.test(df[['mean_sus']], df[[expl]], method='spearman') %>% tidy
  }
}

models = function(df) {
  out = data_frame()
  for (expl in c('cpkp', 'did')) {
    new_rows = safe_spearman(df, expl) %>%
      mutate(explanatory=expl)

@@

res = dat %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(.)) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, state_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

@@

>>>>

```{r run_models}
safe_spearman = function(df, expl, indep) {
  if (dim(df)[1] < 3L) {
    data_frame(estimate=NA, statistic=NA, p.value=NA, method=NA, alternative=NA)
  } else {
    cor.test(df[[indep]], df[[expl]], method='spearman') %>% tidy
  }
}

models = function(df, indep) {
  out = data_frame()
  for (expl in c('cpkp', 'did')) {
    new_rows = safe_spearman(df, expl, indep) %>%
      mutate(explanatory=expl)

@@

res = dat %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, state_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

@@

And I try the same thing again, but this time keeping all the data points:

```{r correlations_with_all_points}
dat %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'percent_susceptible')) %>%
  ungroup %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  filter(!is.na(estimate), state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  select(bug, consumption_drug, resistance_drug,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```
